---
title: "Stat 101C Final Project"
author: "Mizuki Kio"
date: "May 24, 2017"
output: html_document
---

```{r}
setwd("D:/UCLA Document/2016-2017/Spring 2017/Stats 101C/Final Project")

#load data
library(readr)
lafd <- read_csv("D:/UCLA Document/2016-2017/Spring 2017/Stats 101C/Final Project/lafdtraining.csv")
testing <- read_csv("D:/UCLA Document/2016-2017/Spring 2017/Stats 101C/Final Project/testing.without.response.csv")

set.seed(123)
lafd_sample = lafd[sample(1:nrow(lafd), size = 0.1 * nrow(lafd)),]
lafd_sample = na.omit(lafd_sample)
lafd_sample$logElapse = log(lafd_sample$elapsed_time)
lafd_sample = lafd_sample[,-c(1,2,5,11)]

attach(lafd_sample)
```

#simplest model

```{r}
model_simple = lm(logElapse~`Dispatch Status`, data = lafd_sample)
summary(model_simple)

pred = exp( predict(model_simple, newdata = testing) )

myOutput = data.frame(row.id=testing$row.id, prediction = pred)

write.csv(myOutput, file = "myOutput.csv", row.names = FALSE)


```
# MODEL 1: Ver 1.0
I realized that Incident Creation Time is bimodial, maybe day and night helps?
Categorize night as after 18:00 and before 06:00
```{r m1}
is_night <- function(x){
  return (x>=18*3600 || x<=6*3600)
}
lafd_sample$night <- sapply(as.numeric(lafd_sample$`Incident Creation Time (GMT)`),is_night)
testing$night <- sapply(as.numeric(testing$`Incident Creation Time (GMT)`),is_night)


#hist(as.numeric(lafd_sample$`Incident Creation Time (GMT)`))
m1 <- lm(logElapse~as.factor(`First in District`)+`Dispatch Status`+night,data=lafd_sample)
summary(m1)
a <- summary(m1)
sum(a$residuals^2)
```
Summary:
m1: add variable 'night' to indicate whether the incident happens at night or not. 
Night is defined as after 18:00 pm and before 6:00 pm

# MODEL 2: Ver 2.0
Unit Type definitely affect elapsed time. E.g. Ambulances are able to arrive on spot quicker.
```{r m2}
# Learn from the failure previously. Name low frequency unit type as "others"
sort(table(lafd_sample$`Unit Type`),decreasing = TRUE)
# Assign Unit type = "Others" when occurence of that type <10
infreq_type <- names(table(lafd_sample$`Unit Type`))[table(lafd_sample$`Unit Type`)<10]

lafd_sample$type <- lafd_sample$`Unit Type`
lafd_sample$type[lafd_sample$`Unit Type` %in% infreq_type] <-"Others"
sort(table(lafd_sample$type),decreasing = TRUE)


# modify the testing data as well, change the above infrequent types in training as well as types unique to testing data to "Others"
unique_type <- setdiff(testing$`Unit Type`,lafd_sample$`Unit Type`)
unique_type
testing$type <- testing$`Unit Type`
testing$type[testing$`Unit Type` %in% c(infreq_type,unique_type)] <-"Others"
sort(table(testing$type),decreasing = TRUE)


# Fit a model
m2 <- lm(logElapse~as.factor(`First in District`)+`Dispatch Status`+night+type,data=lafd_sample)
summary(m2)
a <- summary(m2)
sum(a$residuals^2)

# prediction
pred <- exp(predict(m2, newdata = testing))

myOutput_m2 = data.frame(row.id=testing$row.id, prediction = pred)

write.csv(myOutput_m2, file = "myOutput_m2.csv", row.names = FALSE)
```
Summary:
m2: add variable night, Unit Type (changed infrequent types as 'others').
Kaggle score: 1788863.93525
